---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# sfcentralities

<!-- badges: start -->
<!-- badges: end -->

sfcentralities allows to find central points of sf objects easily. It offers three different functions to find central points:

st_geo_median is a vectorized algorithm to find the geometric median of a given set of points. The geometric median is the point at which the sum of distances to the other points are minimized.

st_network_median is an algorithm to find the central point within a net, for example, a streetnet. It gives the most central point of a set of points, based on distances along the streets of Openstreetmap  or, if the user prefers it, a streetnet supplied by the users.

st_network_centrality determines closeness centrality of a street network. The most central point in a network is the network node whose sum of distances along the network to all other nodes is the smallest.

The package also offers the function 'st_network_distance'. This is a function which can be used directly to compute distances between two points. It is the basis of st_network_median.

## Installation

You can install the development version of sfcentralities from [GitHub](https://github.com/) with:

``` r
# install.packages("pak")
pak::pak("InaKrapp/sfcentralities")
```

## Examples

## st_geo_median
sffunctions offers the function st_geo_median to calculate geometric medians from datasets with simple features.

The geometric median for a set of points is defined as the point which minimizes the straight-line-distances to all these points. This package uses the same underlying algorithm as the 'pracma' package to calculate it, but the function is vectorized for faster computation. Furthermore, it is designed to use sf objects as input.
```{r}
# Load necessary packages
library(sf)
library(sfcentralities)

# Create example data
pts <- st_sfc(st_point(c(0, 0)), st_point(c(1, 1)), st_point(c(2, 2.2)), st_point(c(-2, 3)))
pts <- st_as_sf(pts)
st_crs(pts) <- "EPSG:3857"

# Calculate the geometric median of the dataset
single_geometric_median <- st_geo_median(pts)
```

A geometric median can be calculated for an entire dataset, or for subgroups within a dataset. To calculate medians for subgroups, the column which defines the groups has to be given as second argument to the function:

```{r}
# Create an index that is 1 for the first three points and 2 for the fourth point:
pts$index <- 1
pts[4,]$index = 2

# Calculate the geometric medians for each index value separately
geometric_medians_by_group <- st_geo_median(pts, "index")

## Look at result:

```
The returned object is a sf object which contains the geometric median as geometry column as well as further information like a column 'number_of_points', the user can check how many points were used to calculate the respective median, and 'distance', which contains the sum of euclidean distances between the points and the geometric median.

Since the result is a sf object, it can be plotted easily with ggplot2:

```{r}
## Plot the results:
library(ggplot2)

ggplot()+geom_sf(data =pts, color = "blue", size = 5 )+geom_sf(data = geometric_medians_by_group, color = "yellow")+ geom_sf(data = single_geometric_median, color = "red")+theme_minimal()
```

Note that there are two groups in the example data: One only contains the point in the upper right, which is therefore also returned as the geometric median of that group.
The second group contains the three points in the lower right. The geometric median is the same as the middle point of them.
Calculating the geometric median of all points, the point in the upper left slightly draws it to the upper left, but the points in the lower right corner are 'dominating', so it is still very close to them.

For instances where the dataset or a subgroup of the dataset has a very low number of points, the geometric median is not defined unambiguously. In such cases, the function will act as follows:
1. If asked to calculate the geometric median from a single point, it will return the point it was given.
2. If asked to calculate the geometric median from two points, it will return the point halfway between them.


## st_closeness_centrality
Within a street network, some points can also be considered more central than others. But unlike on a plane with straight-line distances, there is no unique definition of centrality within a network.

This package implements the closeness centrality, which is defined as the inverse of the sum of distances between a certain point in a network and all other points. In other words, the longer it would take to travel from a point to all other points, the lower is the centrality of this point. 

This package is build on the dodgr-package, which offers highly optimized distance calculations along graphs. It is closely integrated with Openstreetmap, To calculate the closeness centrality for all points within a street network, the following commands can be used:

```{r}
library(dodgr)
# Create an example streetnet from a small section of a city
# Using a bounding box for Cambridge, MA
net <- dodgr_streetnet(bbox = c(-71.12, 42.37, -71.10, 42.38), expand = 0.05)
# Get distances within the streetnet for people who walk on foot.
graph <- weight_streetnet(net, wt_profile = "foot")
# Calculate closeness:
result =  st_closeness_centrality(graph)
```

Alternatively, you can supply a sf object and a mode of transport to the 'st_closeness_centrality'-function:
This is especially intended to be used for cases in which a sf object was obtained using the 'dodgr_streetnet' function or similar cases in which a sf object also represents a network.


Calculating the closeness centrality for a large number of points at once can be computationally demanding. If the street network is larger than a certain size, computations are therefore split up in batches. They will take longer to be completed, but this allows such calculations also for large networks.

### st_network_median
In case the user has a set of points located within a city and wishes to find out which of these points has the highest centrality based on distances along the city's street network, the function 'st_network_median' can be used.
Analog to the idea of the geometric median as the point which minimizes euclidean distances to all points in a set, the network median is defined here as the point which minimizes distances along the streetnet for the given set of points.
For computational reasons (even for a relatively small street network, considering every possible point would quickly bring a computer to its limit) only the points within the set are treated as candidates.

```{r example}
library(sfcentralities)
library(dodgr)
 library(sf)

 # 1. Create a street network graph
 graph <- weight_streetnet(hampi, wt_profile = "foot")

 # 2. Define a set of points of interest
 # For example, three temples in Hampi
 pts_df <- data.frame(
   name = c("Virupaksha Temple", "Vijaya Vittala Temple", "Sasivekalu Ganesha"),
   lat = c(15.3356, 15.3476, 15.3366),
   lon = c(76.4583, 76.4760, 76.4633)
 )
 pts_sf <- st_as_sf(pts_df, coords = c("lon", "lat"), crs = 4326)

 # 3. Find the network median
 median_point <- st_network_median(data = pts_sf, graph = graph)

 # 4. View the result
 print(median_point)
 #> Simple feature collection with 1 feature and 2 fields
 #> Geometry type: POINT
 #> Dimension:     XY
 #> Bounding box:  xmin: 76.46366 ymin: 15.34107 xmax: 76.46366 ymax: 15.34107
 #> Geodetic CRS:  WGS 84
 #>           id sum_of_distances                    geometry
 #> 1 2925345719         4783.336 POINT (76.46366 15.34107)

 # You can plot the result to visualize it
 if (require("ggplot2") && require("ggrepel")) {
  ggplot() +
    geom_sf(data = dodgr_to_sf(graph), color = "gray80") +
   geom_sf(data = pts_sf, color = "blue", size = 3) +
  geom_sf(data = median_point, color = "red", size = 5, shape = 18) +
  geom_text_repel(data = median_point, aes(label = "Central Point",
                                  geometry = geometry),
                  stat = "sf_coordinates", nudge_x = 0.001) +
  theme_void() +
  labs(title = "Most Central Point (Red) Among a Set of Points (Blue)")
 }
```
```{r}
  graph <- weight_streetnet(hampi, wt_profile = "foot")
  pts_df <- data.frame(
    name = c("Virupaksha Temple", "Vijaya Vittala Temple"),
    lat = c(15.3356, 15.3476),
    lon = c(76.4583, 76.4760)
  )
  pts_sf <- st_as_sf(pts_df, coords = c("lon", "lat"), crs = 4326)

  # Action
  median_point <- st_network_median(data = pts_sf, graph = graph)

```
# Change me please.

I should modify the geometry so the geometry of the returned object is not 'x'.
And I should consider modifying the crs since I get a warning saying it may be wrong for long-lat coordinates. Investigate: Where does the warning come from?

```{r}
 library(sf)
 library(dodgr)

# Create an example graph from a small section of a city
# Using a bounding box for Cambridge, MA
 net <- dodgr_streetnet(bbox = c(-71.12, 42.37, -71.10, 42.38), expand = 0.05)
 graph <- weight_streetnet(net, wt_profile = "foot")

# Create a set of random points within that bounding box
# set.seed for reproducibility
 set.seed(123)
 pts <- st_sample(st_as_sfc(st_bbox(c(xmin=-71.12, ymin=42.37, xmax=-71.10, ymax=42.38))), 100)
 pts <- st_as_sf(pts)
 st_crs(pts) <- "EPSG:4326" # Set CRS to WGS84
 pts$id <- 1:nrow(pts) # Add an ID column

# Find the most central point
 central_point <- st_network_median(data = pts, graph = graph)

# Print the result
 print(central_point)

 # You can visualize the result
 # (Requires ggplot2 and ggrepel)
  library(ggplot2)
  library(ggrepel)
 
  ggplot() +
    geom_sf(data = dodgr_to_sf(graph), color = "gray80") +
    geom_sf(data = pts, color = "blue", size = 3) +
    geom_sf(data = central_point, color = "red", size = 5, shape = 18) +
    geom_text_repel(data = central_point, aes(label = "Central Point",
                                     geometry = x),
                    stat = "sf_coordinates", nudge_x = 0.001) +
    theme_void() +
    labs(title = "Most Central Point (Red) Among a Set of Points (Blue)")
```


The test graph is not created correctly yet.
```{r}
# Load necessary libraries for testing
library(testthat)
library(sf)
library(dodgr)

# --- Sourcing the functions to be tested ---
# (Ensure st_network_distance and find_central_point are loaded)

# --- Test Suite ---
# --- 1. Setup a predictable test environment ---

# Create vertices data frame with ID, longitude (x), and latitude (y)
v_linear <- data.frame(
  id = 1:4,
  x = c(0, 1, 2, 3),
  y = rep(0, 4)
)

# Create edges data frame referencing vertex IDs
edges_linear <- data.frame(
  from_id = c(1, 2, 3),
  to_id = c(2, 3, 4)
)

# --- CORRECT METHOD: Construct an sf object with LINESTRINGs ---

# Create a list to hold the LINESTRING geometries
linestrings <- list()
# For each edge, look up the coordinates of its start and end points
# and create a LINESTRING.
for (i in 1:nrow(edges_linear)) {
  from_node <- edges_linear$from_id[i]
  to_node <- edges_linear$to_id[i]
  
  # Get coordinates for the start and end of the line
  from_coords <- v_linear[v_linear$id == from_node, c("x", "y")]
  to_coords <- v_linear[v_linear$id == to_node, c("x", "y")]
  
  # Create a matrix of the two points
  line_matrix <- as.matrix(rbind(from_coords, to_coords))
  
  # Create the LINESTRING and add it to our list
  linestrings[[i]] <- sf::st_linestring(line_matrix)
}

# Convert the list of linestrings into a simple features geometry column (sfc)
sfc_network <- sf::st_sfc(linestrings, crs = 4326)

# Create the final sf object. This is the object weight_streetnet expects.
sf_network <- sf::st_sf(geometry = sfc_network)

# column "highway" and give it a generic value that is walkable.
sf_network$highway <- "path"

# Now, call weight_streetnet with the correct sf object.
# It will correctly process the LINESTRINGs to build a routable graph.
graph_linear <- dodgr::weight_streetnet(sf_network, wt_profile = "foot")

# Create the corresponding sf points for testing the function (this part remains the same)
pts_linear <- sf::st_as_sf(v_linear, coords = c("x", "y"), crs = 4326)
# Add character ID to match what dodgr produces from vertices.
pts_linear$id <- as.character(pts_linear$id)

ggplot()+geom_sf(data = pts_linear)

```

Note that if two points are similarily central in such a network, the function will return only one of them. Due to floating point issues, a warning may not be send.


### st_network_distance

Calculating distance along street networks can offer more accurate results than straight-line distances. The function 'st_network_distance' allows the user to calculate street distances on sf objects using the dodgr package. The st_geo_median-function makes use of this, users can also use it to create their own centrality measures.

The dodgr package is able to calculate distances along networks with high efficiency, taking into account aspects like different modes of transport. It is particularly well-integrated with Openstreetmap, so the typical use case is to let the function calculate distances along the streetnet of Openstreetmap. The user can either supply a streetnet they created before or define a place and a mode of transport. In the latter case, the streetnet will automatically be downloaded from Openstreetmap.

The st_network_distance function can be used in a number of ways:
1. Supplying a single sf object. In this case, it calculates distances between all points in the object:
```{r}
pts <- sf::st_sfc(sf::st_point(c(76.47398, 15.330)), sf::st_point(c(76.47398, 15.150)))
  pts <- sf::st_as_sf(pts)
  sf::st_crs(pts) <- "EPSG:4326"
  # Calculate distance
  result <- st_network_distance(data = pts, placename = "hampi", transport_mode = "bicycle")
```
2. Supplying two sf objects. In this case, the distances from all points of the first object to all points of the second object are calculated.
If the user wants the reverse distances, they can switch the order of the inputs.
It is important to keep in mind that in road networks, the distance from point A to point B can be different from the distance from point B to point A, f.eg. because of one-way streets or construction sites.
```{r}
pts <- sf::st_sfc(sf::st_point(c(76.47398, 15.330)), sf::st_point(c(76.47398, 15.150)))
  pts <- sf::st_as_sf(pts)
  sf::st_crs(pts) <- "EPSG:4326"
  pts2 <- pts[1, ]
  # Calculate distance
  result <- st_network_distance(data = pts, to = pts2, placename = "hampi", transport_mode = "bicycle")
```
The distances are appended as new columns to the first sf object. They contain the distances to the points of the second object in descending order: 'distance_to_1' is the distance to the point of the first row in the second sf object, 'distance_to_2' is the distance to the point of the second row and so on.
For easier oversight, it may make sense to supply names. For example, if the user wants to calculate the distance to two points named 'city center' and 'train_station', they can name the calculated distances after them using the 'index' parameter of the function:
```{r}
pts <- sf::st_sfc(sf::st_point(c(76.47398, 15.330)), sf::st_point(c(76.47398, 15.150)))
  pts <- sf::st_as_sf(pts)
  sf::st_crs(pts) <- "EPSG:4326"
  pts$name = c("city center", "train station")
  # Calculate distance
  result <- st_network_distance(data = pts, to = pts, placename = "hampi", transport_mode = "bicycle", index = "name")
```
A function like this supplies the name of an area ('placename') to Openstreetmap and takes the street network from there. This approach requires a working internet connection, and especially for distances between places which are far apart (f.eg. in different cities), it may fail when it attempts to download too large street networks.

Therefore, it may be more efficient for the user to download the street network and give it directly into the function. This allows, for example, to download it once and save it locally for later uses, or to modify it.
Here is example code to create a graph:
```{r}
  bb <- osmdata::getbb(place_name = "hampi")
  net <- dodgr::dodgr_streetnet(bb)
  graph <- dodgr::weight_streetnet(net, wt_profile = "bicycle")
```
The graph can then be supplied to the st_network_distance function as follows:
```{r}
result <- st_network_distance(data = pts, to = pts, graph = graph)
```
Note that supplying a placename or a mode of transport is no longer needed; when they were supplied previously, they were used to create the graph internally. If a graph is supplied, they are therefore ignored.

Finally, it is possible to create the times instead of the distances by setting 'calculate_times' to True:
```{r}
  result_times <- st_network_distance(data = pts, to = pts, placename = "hampi", transport_mode = "bicycle", calculate_times =  TRUE)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this.
